---
title: "Nolan Ryan Comparison"
author: "Christian Lindke"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing Nolan Ryan's Win/Loss Percentage to the Average Starter on His Teams

Was Nolan Ryan a "Winner"?


To answer this question we need to examine whether his long innings gave him a higher
Winning Percentage than the average STARTER on his team.

We do not want to compare his record to "All Pitchers" on his teams as the W/L criteria
for relievers is different and a lower bar than for starters. Compare Apples to Apples
As much as possible.

First load Ryan's Data with Summary Data for Other Starters

```{r data}
RyanData <- read.csv("https://raw.githubusercontent.com/ChristianLindke/Practice/master/Nolan%20Ryan%20Comma.csv")

```

Let's look at the Head, Tail, and do a Summary.

```{r}
head(RyanData)
tail(RyanData)

summary(RyanData)
```


The Variables of interest here are:

1) W.L.: Nolan Ryan's Win/Loss Percentage
2) Starter.WL.: Starter Win/Loss Percentage for Ryan's Team
3) StarterWL.ExclRyan: The Average Win/Loss Percentage Excluding Ryan

Let's make those.

```{r}
RyanWL <- RyanData$W.L.
RestOfTeamWL <- RyanData$StarterWL.ExclRyan 

```

Now, let's run a T-Test to see if there is a significant difference!

```{r}
t.test(RyanWL, RestOfTeamWL)
```

### Results

That p-value of 0.6307 suggests that there is no sigificant difference between him and his teammates.

Doesn't look good for Nolan Ryan being a significantly better "Winner" than the average of his team mates, but let's visualize this.

AND it looks like his overall winning percentage was WORSE than the average of other starters on his teams.

To visualize the information, we need a couple of additional pieces of information and to put all of that information into a new data frame.

```{r}
MeanRyan <- mean(RyanWL)
SDRyan <- sd(RyanWL)

MeanTeams <- mean(RestOfTeamWL)
SDTeams <- sd(RestOfTeamWL)


Ryan_Career <- mean(RyanWL)
Team_Career <- mean(RestOfTeamWL)

Ryan_Comparison <- data.frame(
  name=c("Ryan Career W/L %", "Rest of Team W/L %"),
  value=c(Ryan_Career, Team_Career),
  sd=c(SDRyan, SDTeams)
)

```


Now that we've got the the Mean W/L percentage for both Ryan and the rest of his team's starters over his career, and the standard deviations of these variables, we can visualize our comparison.

Let's visualize this!

## The Plots

We'll do a basic and a more clean looking plot.

```{r RyanVisualized, echo=FALSE}
library(ggplot2)

ggplot(Ryan_Comparison) +
  geom_bar( aes(x=name, y=value), stat="identity", fill="darkblue", alpha=0.7) +
  geom_errorbar( aes(x=name, ymin=value-sd, ymax=value+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)


ggplot(Ryan_Comparison) +
  geom_bar( aes(x=name, y=value), stat="identity", fill="darkblue", width = 0.25, alpha=0.5) +
  geom_pointrange( aes(x=name, y=value, ymin=value-sd, ymax=value+sd), colour="orange", alpha=0.9, size=1.3)
```

What we see is that Ryan had a slightly lower Win-Loss percentage than his teammates. On it's own, that might be damning to him as a Hall of Famer. Thankfully, those other teammates included other great pitchers like Tom Seaver, Frank Tanana, Joe Niekro, Don Sutton, Mike Scott, Bob Knepper, and Kevin Brown.

```{r, echo=FALSE}

```


As with his career when it came to Cy Young votes, Ryan was frequently the 2nd or 3rd best pitcher on his own team. He was a great pitcher, who's longevity, power, and ability to adapt warrant him being in the Hall, but he is also an example of how his legacy sometimes gets us to forget some of the greats he played with.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
